\clearpage
\item \points{20} {\bf Bayesian Interpretation of Regularization}

\textbf{Background: }
In Bayesian statistics, almost every quantity is a random variable, which
can either be observed or unobserved. For instance, parameters $\theta$ are
generally unobserved random variables, and data $x$ and $y$ are observed
random variables. The joint distribution of all the random variables is
also called the \emph{model} (\emph{e.g.,} $p(x, y, \theta)$). Every unknown quantity can
be estimated by conditioning the model on all the observed quantities. Such
a conditional distribution over the unobserved random variables, conditioned
on the observed random variables, is called the \emph{posterior distribution}.
For instance $p(\theta | x, y)$ is the posterior distribution in the
machine learning context. A consequence of this approach is that we are
required to endow our model parameters, \emph{i.e.,} $p(\theta)$, with a \emph{prior distribution}.
The prior probabilities are to be assigned \emph{before} we see the data---they capture our prior beliefs of what the model parameters might be
before observing any evidence.

In the purest Bayesian interpretation, we are required to keep the entire
posterior distribution over the parameters all the way until prediction, to
come up with the \emph{posterior predictive distribution}, and the final prediction
will be the expected value of the posterior predictive distribution. However
in most situations, this is computationally very expensive, and we settle for
a compromise that is \emph{less pure} (in the Bayesian sense).

The compromise is to estimate a point value of the parameters (instead of the
full distribution) which is the mode of the posterior distribution. Estimating
the mode of the posterior distribution is also called
\emph{maximum a posteriori estimation} (MAP). That is, 
$$\theta_{\text{MAP}} = \arg\max_\theta p(\theta|x,y).$$
Compare this to the \emph{maximum likelihood estimation} (MLE) we have
seen previously:
$$\theta_{\text{MLE}} = \arg\max_\theta p(y|x,\theta).$$
In this problem, we explore the connection between MAP estimation, and common
regularization techniques that are applied with MLE estimation.
In particular, you will show how the choice of prior distribution over $\theta$ (\emph{e.g.,}
Gaussian or Laplace prior)
is equivalent to different kinds of regularization (\emph{e.g.,} $L_2$, or $L_1$
regularization). To show this, we shall proceed step by step, showing intermediate
steps.

\begin{enumerate}
    \input{03-bayesian-regularization/01-argmax}
    \input{03-bayesian-regularization/02-l2}
    \input{03-bayesian-regularization/03-closed-form}
    \input{03-bayesian-regularization/04-l1}   
\end{enumerate}

\textbf{Remark:} Linear regression with $L_2$ regularization is also commonly called \emph{Ridge regression}, and when $L_1$ regularization is employed, is commonly called \emph{Lasso regression}. These regularizations can be applied to any Generalized Linear models just as above (by replacing $\log p(y|x,\theta)$ with the appropriate family likelihood). Regularization techniques of the above type are also called \emph{weight decay}, and \emph{shrinkage}. The Gaussian and Laplace priors encourage the parameter values to be closer to their mean (\emph{i.e.,} zero), which results in the shrinkage effect.

\textbf{Remark:} Lasso regression (\emph{i.e.,} $L_1$ regularization) is known to result in sparse parameters, where most of the parameter values are zero, with only some of them non-zero.
